FraCaS: Temporal Analysis

### 2020-06-09 paper2020/paper.tex
wibble
import example

\section{Our Treatment of time}
\todo{JP: add a completely worked out example}
\label{sec:our-semantics}

@@

\[\begin{array}{l}
∃x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\]
Note here that the scope for the existential is extended beyond the
scope of \textbf{P1}. This extension can work according to
\citet{unger_dynamic_2011}, and our implemented analysis of anaphora\ifanon (citation anonymized) \else\citep{bernardy_computational_2018,bernardy_wide_2019}\fi.

@@

\[\begin{array}{l}
∃x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃y. novel(y) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\]
Our analysis does not need to treat this last case specially. Indeed,
even if $write(smith,x,.,.)$ is an activity and thus subject to
unicity of action, in (280), $x$ is quantified existentially; we have
two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot
deduce equality of the intervals $t_1,t_2$ and $t_3,t_4$. In turn, the
hypothesis cannot be refuted.

@@

>>>>

\section{Our Treatment of time}
\label{sec:our-semantics}

@@

\[\begin{array}{l}
∀x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\]
Note here that the scope for the existential is extended beyond the
scope of \textbf{P1}, and its polarity switched (to universal). This extension can work according to
\citet{unger_dynamic_2011}, and our implemented analysis of
anaphora\ifanon (citation anonymized)
\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.

@@

\[\begin{array}{l}
(∃x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2)) ∧ \\
(∃y. novel(y) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4)) \\
⟶ ⊥ \\
\end{array}
\]
Our analysis does not need to treat this last case specially. Indeed,
even if $write(smith,x,.,.)$ is an activity and thus subject to
unicity of action, in (280), $x$ is quantified existentially; we have
two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot
deduce equality of the intervals $t_1,t_2$ and $t_3,t_4$. In turn, the
hypothesis cannot be refuted.

@@

\section{Worked out example}
To give a sense of the additional details necessary to deal with the
precision demanded by a proof-assistant such as Coq we show how
problem (279) is worked out fully.

Input from \citet{Ljunglof:2012}:

\begin{verbatim}
s_279_1_p :: Phr
s_279_1_p=((sentence) ((useCl) ((past)) ((pPos)) ((predVP) ((usePN) ((lexemePN "smith_PN"))) ((advVP) ((complSlash) ((slashV2a) ((lexemeV2 "write_V2"))) ((detCN) ((detQuant) ((indefArt)) ((numSg))) ((useN) ((lexemeN "novel_N"))))) ((lexemeAdv "in_1991_Adv"))))))
s_279_3_h :: Phr
s_279_3_h=((sentence) ((useCl) ((past)) ((pPos)) ((predVP) ((usePN) ((lexemePN "smith_PN"))) ((advVP) ((complSlash) ((slashV2a) ((lexemeV2 "write_V2"))) ((usePron) ((it_Pron)))) ((lexemeAdv "in_1992_Adv"))))))
\end{verbatim}
Interpretation by our dynamic and temporal semantics:

\begin{verbatim}
Definition Problem279aFalse:= (FORALL (fun a=>novel_N a) (fun a=>(exists (b: Time), ((exists (c: Time), (IS_INTERVAL Date_19910101 b /\ IS_INTERVAL c Date_19911231 /\ IS_INTERVAL b c /\ appTime b c (write_V2 a) (PN2object smith_PN))))) -> Not (exists (f: Time), ((exists (g: Time), (IS_INTERVAL Date_19920101 f /\ IS_INTERVAL g Date_19921231 /\ IS_INTERVAL f g /\ appTime f g (write_V2 a) (PN2object smith_PN))))))).
\end{verbatim}
Unfolding definitions by Coq:

\begin{verbatim}
  forall x : object,
  novel_N x ->
  (exists b c : Z,
     Date_19910101 <= b /\ c <= Date_19911231 /\ b <= c /\ write_V2 x SMITH b c) ->
  (exists f g : Z,
     Date_19920101 <= f /\ g <= Date_19921231 /\ f <= g /\ write_V2 x SMITH f g) ->
  False
\end{verbatim}
Proof:

\begin{verbatim}
Theorem  problem279 : Problem279aFalse.
cbv.
intros novel isSmithsNovel P1 H.
destruct P1 as [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
destruct H as [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
specialize writeUnique with (x := novel)(y := SMITH) as A.
unfold UniqueActivity in A.
specialize (A _ _ _ _ P1 H) as B.
lia.
Qed.
\end{verbatim}
